%---------------------------------------------------------------%
An urn contains 10 balls: 4 red and 6 blue. A second urn contains 16 red balls and an 
unknown number of blue balls. A single ball is drawn from each urn. The probability 
that both balls are the same color is 0.44 . 
Calculate the number of blue balls in the second urn. 
 (A) 4 
 (B) 20 
 (C) 24 
 (D) 44 
 (E) 64

%----------------------------------------------------------------%


An auto insurance company has 10,000 policyholders. Each policyholder is classified as 
 (i) young or old; 
 (ii) male or female; and 
 (iii) married or single. 
Of these policyholders, 3000 are young, 4600 are male, and 7000 are married. The 
policyholders can also be classified as 1320 young males, 3010 married males, and 1400 
young married persons. Finally, 600 of the policyholders are young married males. 
 How many of the company’s policyholders are young, female, and single? 

(A) 280 
(B) 423 
(C) 486 
 (D) 880 
(E) 896 

%----------------------------------------------------------------%

In a class of 20 pupils, 11 have dark hair, 7 have fair hair and 2 have red hair.

Two pupils are chosen at random to collect the homework. What is the probability that they

a) both have fair hair

(3 marks)

b) each have hair of a different colour

(4 marks)

(Marks available: 7)

Answer outline and marking scheme for question: 2

a) 21/190

(3 marks)

b) 226/380

(4 marks)
%----------------------------------------------------------------%
