\documentclass[a4]{beamer}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{newlfont}
\usepackage{amsmath,amsthm,amsfonts}
%\usepackage{beamerthemesplit}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{mathptmx}  % Font Family
\usepackage{helvet}   % Font Family
\usepackage{color}
\mode<presentation> {
 \usetheme{Default} % was
 \useinnertheme{rounded}
 \useoutertheme{infolines}
 \usefonttheme{serif}
 %\usecolortheme{wolverine}
% \usecolortheme{rose}
\usefonttheme{structurebold}
}
\setbeamercovered{dynamic}

\title[MA4413]{Statistics for Computing \\ {\normalsize MA4413 Lecture 4A and 4B}}
\author[Kevin O'Brien]{Kevin O'Brien \\ {\scriptsize Kevin.obrien@ul.ie}}
\date{Autumn Semester 2013}
\institute[Maths \& Stats]{Dept. of Mathematics \& Statistics, \\ University \textit{of} Limerick}

\renewcommand{\arraystretch}{1.5}
\begin{document}

\begin{frame}
\titlepage
\end{frame}
%---------------------------------------------------------------------%
\frame{
\frametitle{Week 4 (from Lecture 4A)}
\begin{itemize}
\item The first midterm is to take place Monday of Week 5 at 4pm.
\item The first midterm will cover:
\begin{itemize}
\item Basic Probability
\item Descriptive statistics (mean, median variance etc)
\item Discrete probability distributions (binomial and Poisson)
\item The exponential distribution
\item Some of the normal distribution will be included.
\end{itemize}
\end{itemize}
}

\begin{frame}[fragile]
\frametitle{Continuous Distributions: Current Status}
\begin{itemize}
\item (The Continuous Uniform Distribution, Not examinable)
\item The Exponential Distribution (Examinable for midterm)
\item (Exponential Distribution is the Cut-off point for Mid-Term 1)
\item The Normal Distribution
\item The Standard Normal (Z) Distribution.
\item Applications of Normal Distribution
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Exponential Distribution}
The Exponential Distribution may be used to answer the following questions:
\begin{itemize}
\item How much time will elapse before an earthquake occurs in a given region?
\item How long do we need to wait before a customer enters our shop?
\item How long will it take before a call center receives the next phone call?
\item How long will a piece of machinery work without breaking down?
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Exponential Distribution}

\begin{itemize}
\item All these questions concern the time we need to wait before a given event occurs. If this waiting time is unknown, it is often appropriate to think of it as a random variable having an exponential distribution.
\item Roughly speaking, the time $X$ we need to wait before an event occurs has an exponential distribution if the probability that the event occurs during a certain time interval is proportional to the length of that time interval.

\end{itemize}
\end{frame}

%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Probability density function}
The probability density function (PDF) of an exponential distribution is

\[
f(x) = \begin{cases}
\lambda e^{-\lambda x}, & x \ge 0, \\
0, & x < 0.
\end{cases}\]
The parameter $\lambda$  is called \textbf{\emph{rate}} parameter.
\end{frame}

%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Cumulative density function}
The cumulative distribution function (CDF) of an exponential distribution is

\[
P(X \leq x) = F(x) = \begin{cases}
1-e^{-\lambda x}, & x \ge 0, \\
0, & x < 0.
\end{cases}\]


The complemeent of the CDF (i.e. $P(X \geq x)$ is

\[
P(X \geq x) = \begin{cases}
e^{-\lambda x}, & x \ge 0, \\
0, & x < 0.
\end{cases}\]
\end{frame}

%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Expected Value and Variance}
The expected value of an exponential random variable $X$ is:

\[
E[X] = \frac{1}{\lambda}\]
The variance of an exponential random variable $X$ is:

\[
V[X] = \frac{1}{\lambda^2}\]

\end{frame}

%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Exponential Distribution: Example}
Assume that the length of a phone call in minutes is an exponential random variable $X$ with parameter
$\lambda = 1/10$. If someone arrives at a phone booth just before you arrive, find the probability that you
will have to wait \begin{itemize}
\item[(a)] less than 5 minutes,
\item[(b)] between 5 and 10 minutes.
\end{itemize}
\end{frame}



%%------------------------------------------------------------------------%
%\begin{frame}[fragile]
%\frametitle{Exponential Distribution: Example}
%\begin{verbatim}
%> dexp(0:10,rate=0.10)
% [1] 0.10000000 0.09048374 0.08187308 0.07408182 0.06703200 0.06065307
% [7] 0.05488116 0.04965853 0.04493290 0.04065697 0.03678794
%>
%> pexp(0:10,rate=0.10)
% [1] 0.00000000 0.09516258 0.18126925 0.25918178 0.32967995 0.39346934
% [7] 0.45118836 0.50341470 0.55067104 0.59343034 0.63212056
%\end{verbatim}
%\end{frame}

%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Exponential Distribution: Example}

\begin{itemize}
\item[(a)] $P(X \leq 5)$ = 0.39346934
\item[(b)] $P(5 \leq X \leq 10)$ \\ = $P( X \leq 10) - P( X \leq 5)$ \\ = 0.6321- 0.3934 \\ = 0.2386 \\= 23.86 $\%$
\item[(c)] Alternative approach to (b)\\$P(5 \leq X \leq 10)$ \\ = $P( X \geq 5) - P( X \geq 10)$ \\
= $e^{-0.5} - e^{-1}$
=0.6065 - 0.3678\\
= 0.2386 = 23.86 $\%$
\end{itemize}

\end{frame}



%------------------------------------------------------------------------%
\begin{frame}[fragile]
\frametitle{Exponential Distribution}
\begin{itemize}
\item The Exponential Rate
\item Related to the Poisson mean (m)
\item If we expect 12 occurrences per hour - what is the rate?
\item We would expected to wait 5 minutes between occurrences.
\end{itemize}
\end{frame}

\end{document} 