
\documentclass{report}
\usepackage{natbib}
\usepackage{vmargin}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{subfigure}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm, amsmath}
%\usepackage[dvips]{graphicx}
\bibliographystyle{chicago}
\renewcommand{\baselinestretch}{1.1}

\begin{document}
\begin{enumerate}
\item


%--descriptive statistics
The mean height of the women in a large population is 1.671m while the mean height
of the men in the population is 1.758m. The mean height of all the members of the
population is 1.712m. Calculate the percentage of the population who are women.




%--- Chi Squared

\item An analyst of the retail trade uses as analytical tools the concepts of ‘Footfall’ (the daily
number of customers per unit sales area of a shop) and ‘Ticket Price’ (the average sale
price of an item in the shop’s offer).
\\
Shops are classified as offering Low, Medium or High price items and, during any sales
period, as having Low, Medium or High footfall.
\\
During the January Sales the analyst studies a sample of shops and obtains the
following frequency data for the nine possible combined classifications:

\begin{tabular}{|c|c|c|c|}
  \hline
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
   & Low Price  & Medium Price & High Price \\\hline
  Low Footfall & 45 & 75 & 23 \\ \hline
  Medium Footfall & 37 & 126 & 25 \\ \hline
  High Footfall & 22 & 43 & 16 \\
  \hline
\end{tabular}
\\
\\
Conduct a suitable test for association between Ticket classification and Footfall level,
and report on your findings.

\end{enumerate}



%-----------------------------------------------------------------------------------------------------------%
\newpage
\section*{Probability Distributions}


\begin{itemize}
\item The discrete uniform distribution
\item The continuous uniform distribution
\item The binomial disribution
\item The poisson distribution
\item The exponential distribution
\item the Normal distribution
\end{itemize}



\begin{enumerate}
%--Distributions




\item uniform - The lower and upper bounds are 13cm and 21cm respectively.

\item A computer server breaks down, on average, once every three months.
\begin{itemize}
\item What is the probability that the server breaks down three times in a quarter?
\item What is the probability that a server breaks down exactly five times in one year?
\end{itemize}

\item Assume that the number of weekly study hours for students at a certain university
is approximately normally distributed with a mean of 22 and a standard deviation
of 6.
\begin{itemize}
\item Find the probability that a randomly chosen student studies less than 12
hours.
\item Estimate the percentage of students that study more than 37 hours.
\end{itemize}

\item
A charity believes that when it puts out an appeal for charitable donations the
donations it receives will normally distributed with a mean €50 and standard
deviation €6, and it is assumed that donations will be independent of each
other.
\begin{itemize} \item Find the probability that the first donation it receives will be less than
€40.
\item Find the probability that it will be between €40 and €45.
\item Find the value $x$ such that 5\% of donations are more than €$x$.
\end{itemize}

\item In an examination the scores of students who attend schools of type A are
normally distributed about a mean of 55 with a standard deviation of 6. The
scores of students who attend type B schools are normally distributed about a
mean of 60 with a standard deviation of 5.
Which type of school would have a higher proportion of students with marks above 70?
\end{enumerate}

%-----------------------------------------------------------------------------------------------------------%
\newpage
\section*{Inference}
\begin{enumerate}
%-- hypothesis testing

\item In the past, 18\% of shoppers have bought a particular brand of breakfast cereal.
After an advertising campaign, a random sample of 220 shoppers is taken and 55 of the sample have bought this brand of cereal.
\begin{itemize}
\item Write down the null and the alternative hypothesis for this problem
\item State whether it is a one tailed or two tailed test
\end{itemize}


%---

\item The starting annual salaries for students graduating from two departments, $X$ and
$Y$, of a university are being investigated. Two random samples of last year’s intake
have been selected and the results are as follows:



\begin{tabular}{|c|c|c|c|}
  \hline
  Dept & Sample size & Mean starting
 salary & Std. Deviation \\\hline
  X & 50 & €21,000 &  €4,900\\\hline
  Y & 40 & €20,000 & €3,000 \\
  \hline
\end{tabular}


\begin{itemize}
\item What proportion of new graduates from Department X earn more than €22,000
per month?
\item Test the hypothesis that graduates from Department X earn more than those from
Department Y at two appropriate levels and comment on your results. Give any
necessary conditions for your test to be valid.
\end{itemize}

%---
\item  A market research company has conducted a survey of adults in two large towns, either
side of an international border, in order to judge attitudes towards a controversial
internationally broadcast celebrity television programme.
The following table shows some of the information obtained by the survey:

 \begin{tabular}{|c|c|c|}
   \hline
   % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    & Town A & Town Z \\ \hline
   Sample size & 50 & 50 \\ \hline
   Sample number approving & 26 & 22 \\
   \hline
 \end{tabular}

\begin{itemize}
\item Conduct a formal hypothesis test, at the 5\% significance level, of the claim that the
population proportions approving the programme in the two towns are equal.

\item Would your conclusion be the same if, in both towns the sample sizes had been 100
(with the same sample proportions of approvals)?
\end{itemize}

%---

\item The intelligence quotient (IQ) of 36 randomly chosen students was measured.
Their average IQ was 109.9 with a variance of 324.
The average IQ of the population as a whole is 100.

\begin{itemize}
\item Calculate the p-value for the test of the hypothesis that on average
students are as intelligent as the population as a whole against the alternative that on average students are more intelligent.


\item Can we conclude at a significance level of 1\% that students are on average more intelligent than the population as a whole?

\item Calculate a 95\% confidence interval for the mean IQ of all students.

\end{itemize}

%---

\item The manufacturer of the new spray also claims that it can be used to
prevent the loss due to insect damage of tender seedlings. To test this
claim, the grower sprays 50 tomato seedlings with the new spray and his
remaining 100 tomato seedlings with his standard spray. After six weeks,
the fruit grower counts the number of healthy plants with the following
results.


\begin{tabular}{|c|c|c|}
  \hline
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
   & New spray & Standard spray \\\hline
  No. of seedlings
sprayed & 50 & 100 \\\hline
  No. of healthy plants
at 6 weeks & 40 & 70 \\
  \hline
\end{tabular}
\\
\\
Construct an approximate 95\% confidence interval for the difference in
the proportion of healthy plants six weeks after spraying between the two
groups.

\end{enumerate}

\end{document}