\documentclass[a4]{beamer}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{newlfont}
\usepackage{amsmath,amsthm,amsfonts}
%\usepackage{beamerthemesplit}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{mathptmx}  % Font Family
\usepackage{helvet}   % Font Family
\usepackage{color}
\mode<presentation> {
 \usetheme{Default} % was
 \useinnertheme{rounded}
 \useoutertheme{infolines}
 \usefonttheme{serif}
 %\usecolortheme{wolverine}
% \usecolortheme{rose}
\usefonttheme{structurebold}
}
\setbeamercovered{dynamic}
\title[MA4413]{Statistics for Computing \\ {\normalsize Lecture 1B}}
\author[Kevin O'Brien]{Kevin O'Brien \\ {\scriptsize Kevin.obrien@ul.ie}}
\date{Autumn Semester 2011}
\institute[Maths \& Stats]{Dept. of Mathematics \& Statistics, \\ University \textit{of} Limerick}
\renewcommand{\arraystretch}{1.5}
\begin{document}



%------------------------------------------------------------%
%-----------------------------------------------------------------%
\frame{
\frametitle{Last lecture}
In the last lecture we looked at how to compute
\begin{itemize}
\item the expected value
\item the variance 
\end{itemize}
of a discrete random variable. In our example, we considered the experiment of throwing a fair die.

}

%-----------------------------------------------------------------%
\frame{
\frametitle{Crooked die}
\begin{itemize}
\item Consider the random experiment of rolling a `crooked' six-sided die, i.e. the outcomes of the throw occur with different probabilities.

\item Suppose we have a die with which an outcome `5' or `6' is twice as likely to occur compared to the other numbers.

\item What is the probability of each outcome?

\begin{itemize}  \item Remark: The ratio of outcomes is 1:1:1:1:2:2
\end{itemize}

\item The probability distributiion can be tabulated as follows

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
$x_i$  & 1 & 2 & 3 & 4 & 5 & 6 \\\hline
$p(x_i)$ & 1/8 & 1/8& 1/8 & 1/8& 2/8 & 2/8\\
\hline
\end{tabular}
\end{center}

\item What is expected value and variance of the outcomes?
\end{itemize}
}

%------------------------------------------------------------------%
\frame{
\frametitle{Variance of the crooked die}

Recall the formula for computing the variance of a discrete random variable:

\[ V(x) = E(X^2) - E(X)^2 \]

We must compute $E(X^2)$

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
$x_i$  & 1 & 2 & 3 & 4 & 5 & 6 \\\hline
$x^2_i$  & 1 & 4 & 9 & 16 & 25 & 36 \\\hline
$p(x_i)$ & 1/8 & 1/8& 1/8 & 1/8& 2/8 & 2/8\\
\hline
\end{tabular}
\end{center}

\[E(X) = (0 \times 1/8) + (1 \times 1/8) +  \ldots + (25 \times 2/8) + (36 \times 2/8) = {32 \over 8} = 4\]

}
%------------------------------------------------------------------%
\frame{
\frametitle{Expected value of the crooked die}

What is the variance?


}

\end{document}
%---------------------------------------------------------------------------%
\frame{
\frametitle{Today's Class}
\begin{itemize}
\item Discrete Probability Distributions
\item Probability Mass Function
\item The Binomial Distribution
\item The Poisson Probability Distribution
\end{itemize}
}

%---------------------------------------------------------------------------%
\frame{
\frametitle{Probability Mass Function}
\begin{itemize} \item a probability mass function (pmf) is a function that gives the probability that a discrete random variable is exactly equal to some value. \item The probability mass function is often the primary means of defining a discrete probability distribution \end{itemize}
}
%---------------------------------------------------------------------------%


\frame{
\frametitle{The Binomial Probability Distribution}
\begin{itemize}
\item The number of independent trials is denoted $n$.
\item The probability of a `success' is $p$
\item The expected number of `successes' from $n$ trials is $E(X) = np$
\end{itemize}
}


\frame{
\frametitle{Binomial Experiment}
A binomial experiment (also known as a Bernoulli trial) is a statistical experiment that has the following properties:
\begin{itemize}
\item The experiment consists of n repeated trials.
\item Each trial can result in just two possible outcomes. We call one of these outcomes a success and the other, a failure.
\item The probability of success, denoted by P, is the same on every trial.
\item The trials are independent; that is, the outcome on one trial does not affect the outcome on other trials.
\end{itemize}
}
%--------------------------------------------------------------------------------------%
\frame{
Consider the following statistical experiment. You flip a coin 2 times and count the number of times the coin lands on heads. This is a binomial experiment because:
\begin{itemize}
\item The experiment consists of repeated trials. We flip a coin 2 times.
\item Each trial can result in just two possible outcomes - heads or tails.
\item The probability of success is constant - 0.5 on every trial.
\item The trials are independent; that is, getting heads on one trial does not affect whether we get heads on other trials.
\end{itemize}
}

%--------------------------------------------------------------------------------------%
\frame{
\frametitle{ Binomial Example 1}

Suppose a die is tossed 5 times. What is the probability of getting exactly 2 fours?

Solution: This is a binomial experiment in which the number of trials is equal to 5, the number of successes is equal to 2, and the probability of success on a single trial is 1/6 or about 0.167. Therefore, the binomial probability is:

\[P(X=2) = ^5C_2 \times (1/6)^2 \times (5/6)^3 = 0.161\]
}


%---------------------------------------------------------------------------%
\frame{
\frametitle{Characteristics of a Poisson Experiment}
A Poisson experiment is a statistical experiment that has the following properties:
\begin{itemize}
\item The experiment results in outcomes that can be classified as successes or failures.
\item The average number of successes (m) that occurs in a specified region is known.
\item The probability that a success will occur is proportional to the size of the region.
\item The probability that a success will occur in an extremely small region is virtually zero.
\end{itemize}
Note that the specified region could take many forms. For instance, it could be a length, an area, a volume, a period of time, etc.
}

%---------------------------------------------------------------------------%
\frame{
\frametitle{Poisson Distribution}
A Poisson random variable is the number of successes that result from a Poisson experiment.
 
The probability distribution of a Poisson random variable is called a Poisson distribution.

Given the mean number of successes ($m$) that occur in a specified region, we can compute the Poisson probability based on the following formula:
}

%---------------------------------------------------------------------------%
\frame{
\frametitle{The Poisson Probability Distribution}
\begin{itemize}
\item The number of occurrences in a unit period (or space)
\item The expected number of occurrences is $m$
\end{itemize}
}

%---------------------------------------------------------------------------%
\frame{
\frametitle{Poisson Formulae}
The probability that there will be $k$ occurences in a unit time period is denoted $P(X=k)$, and is computed as follows.
\Large
\[ P(X = k)=\frac{m^k e^{-m}}{k!} \]

}
%---------------------------------------------------------------------------%
\frame{
\frametitle{Poisson Formulae}
Given that there is on average 2 occurences per hour, what is the probability of no occurences in the next hour? \\ i.e. Compute $P(X=0)$ given that $m=2$
\Large
\[ P(X = 0)=\frac{2^0 e^{-2}}{0!} \]
\normalsize
\begin{itemize}
\item $2^0$ = 1
\item $0!$ = 1
\end{itemize}
The equation reduces to 
\[ P(X = 0)=e^{-2} = 0.1353\]
}
%---------------------------------------------------------------------------%
\frame{
\frametitle{Poisson Formulae}
What is the probability of one occurences in the next hour? \\ i.e. Compute $P(X=1)$ given that $m=2$
\Large
\[ P(X = 1)=\frac{2^1 e^{-2}}{1!} \]
\normalsize
\begin{itemize}
\item $2^1$ = 2
\item $1!$ = 1
\end{itemize}
The equation reduces to
\[ P(X = 1) = 2 \times e^{-2} = 0.2706\]
}
%---------------------------------------------------------------------------%

\end{document} 